Reproducible Research: Peer Assessment 1
=======================================================
## Loading libraries
```{r}
# install.packages("dplyr")
# install.packages("knitr")
# install.packages("tidyr")
# install.packages("lattice")
library(dplyr)
library(knitr)
library(tidyr)
library(lattice)
```

## Loading and preprocessing the data
```{r}
dataframe <- read.csv("activity.csv")
data_tbl <- tbl_df(dataframe)
data_tbl <- filter(data_tbl, !is.na(steps))
data_tbl
```

## Calculate the mean total number of steps taken per day
```{r}
by_date <- group_by(data_tbl, date)
summary_date <- summarize(by_date, steps = sum(steps))
summary_date
hist(summary_date$steps, breaks = 10, col = "green", xlab = "steps", ylab = "frequency",
     main = "Histogram of number of steps taken per day")
rug(summary_date$steps)
abline(v = median(summary_date$steps), lwd = 4, col = "blue")
median(summary_date$steps)
mean(summary_date$steps)

```
## What is the average daily activity pattern?
```{r}
by_interval <- group_by(data_tbl, interval)
summary_interval <- summarize(by_interval, steps = mean(steps))
summary_interval
plot(summary_interval$interval, summary_interval$steps, type = "l", xlab = "interval", ylab = "steps",
     main = "Time series plot of steps") 
# The time interval with the averaged max steps
summary_interval$interval[summary_interval$steps == max(summary_interval$steps)]
```
## Imputing missing values
```{r} 
sum(is.na(dataframe$steps))
# Fill in missing values in a new dataframe(according to the mean of each time interval)
missing_indices <- which(is.na(dataframe$steps), arr.ind=TRUE)
targetIntervals <- dataframe$interval[missing_indices]
fillIn <- c();
for(i in targetIntervals)
  fillIn <- c(fillIn, summary_interval$steps[summary_interval$interval == i]);
newDataframe <- dataframe
newDataframe$steps[is.na(dataframe$steps)] <- fillIn

data_tbl_new <- tbl_df(newDataframe)
by_date_new <- group_by(data_tbl_new, date)
summary_date_new <- summarize(by_date_new, steps = sum(steps))
summary_date_new
hist(summary_date_new$steps, breaks = 10, col = "green", xlab = "steps", ylab = "frequency",
     main = "Histogram of number of steps taken per day(after filling NA's)")
rug(summary_date_new$steps)
abline(v = median(summary_date_new$steps), lwd = 4, col = "blue")
median(summary_date_new$steps)
mean(summary_date_new$steps)
```
After imputing the NA's, the mean stays the same(because we fill NA's with mean step values of corresponding intervals), which cause them to be non-integer. So the median would be non-integer, too.

## Are there differences in activity patterns between weekdays and weekends?
```{r} 
wday <- as.POSIXlt(summary_date_new$date)$wday > 0 & as.POSIXlt(summary_date_new$date)$wday < 6
newDataframe$dayType <- factor(wday, labels = c("weekend", "weekday"))
data_tbl_new <- tbl_df(newDataframe)
by_interval_new <- group_by(data_tbl_new, interval, dayType)
summary_interval_new <- summarize(by_interval_new, steps = mean(steps))
summary_interval_new
xyplot(summary_interval_new$steps ~ summary_interval_new$interval 
       | summary_interval_new$dayType, layout = c(1, 2),
       type = "l", xlab = "Interval", ylab = "Number of steps")
  
```
